GEE models

### 2017-07-20 doc/submission/reviewer-response.md

    We have included a sensitivity analysis identifying which exact covariate
    attenuates the results between the unadjusted and adjusted models; the
    covariate is waist circumference. While inclusion of this variable
    attenuates most of the associations, we believe that adjustment of this
    variable is necessary given that biologically NEFA come from the adipose
    tissue during fasting and hypothetically more adipose tissue would indicate
    higher NEFA. Since other aspects of adipose tissue (e.g. role in
    inflammation and appetite regulation) could also impact insulin sensitivity
    and beta-cell function, we believe it is an important confounder to adjust
    for. We have included in the conclusions a discussion on how adjustment of
    waist circumference influences the model and the interpretation.

5. *"INTRODUCTION: a clear and well-written introduction with a clear objective
and hypothesis"*

6. *"MATERIALS & METHODS:"*
    - *"ESM Figure 1 is not clear to me: a) Prevalent diabetes cases were excluded
    at baseline (n=XX); b) the lines suggest that everyone had to have a year 6
    visit, ehile in the main text, the authors state that any follow-up visit
    was enough for being included (this corresponds well with the GEE
    requirements); c) what happened with incident diabetes cases? Were treated
    diabetes cases included in the analysis? How does that effect measures of
    insulin sensitivity and beta-cell function? I suggest to exclude those
    visits where diabetes was treated."*

        We have clarified how we dealt with incident diabetes cases. Briefly, we
        only report incident diabetes and dysglycemia cases in the text to
        describe the cohort, but in statistical analyses and modeling, these
        cases were excluded. Regarding a) within the main PROMISE cohort,
        prevalent diabetes cases were not excluded from the study as they
        continued to be followed; prevalent diabetes cases did not have fatty
        acids measured. For b) any participant is followed until they request to
        be removed from the study; some participants missed the 3-yr visit, but
        came in for the 6-yr visit. We have included wording within the results
        on the sample and percentage who attended at least two visits. Lastly,
        for c) because diabetes can effect the values of the OGTT-derived 
        indices due to medication the patient is taking and due to diabetes 
        itself impacting beta-cell function and insulin resistance, we excluded 
        them from all analyses; often, these participants did not even complete 
        an OGTT since it is discouraged that these patients consume that much 
        glucose. To restate, diabetes cases were not included in GEE or other 
        analyses.

### 2017-07-20 doc/submission/reviewer-response.md

    - *"I am not an expert of PLS-DA but it is not clear to me, why does this
    analysis reflect more the role of relative NEFA (mole%) analysis than the
    concentration analysis? As I can see, it depends on what variables you feed
    in the model."*

        We are not entirely clear on the reviewer's comment. PLS extracts from
        high dimensional data (multiple predictor variables, i.e. the fatty
        acids) underlying correlations, constrained by a outcome variable (i.e.
        beta-cell function). In this context, because mol% data is by definition
        restricted to sums of 100% (all fatty acid values add up to 100%), the
        underlying correlations between fatty acids is dependent on the relative
        contributions those fatty acids have to the total and as such can
        reflect correlations of specific fatty acids or groups of fatty acids
        that increase or decrease in proportion together. The concentration data
        is not contrained by this inherent correlation and thus underlying groups 
        or clusters may not be identified using PLS.

7. *"RESULTS:"*
    - *"Please give ns in Table 1."* ... *"It would be also nice to see, how
    many participants had 1, 2 or 3 time points in the analysis."*

        These have been added as requested. {{Tony, not sure about how I feel
        adding p-values to Table 1. I'm showing it purely for descriptive
        purposes and it might detract from discussion since, as you may recall,
        some variables (e.g. HDL) have a significant difference between visits,
        but don't really change over the visits... thoughts? I'm not too keen on
        generating p-values for the sake of generating p-values, esp. when it is
        not really relevant to the main analysis.}}

### 2017-07-25 doc/esm-figures.Rmd
generated gee results with no waist, added to esm

sfig_consort <- sfig('consort', 'CONSORT diagram of sample size at each examination visit.')
sfig_dag_isi <- sfig("dag_isi", "Directed acyclic graph taken from DAGitty (DAGitty.net) for the potential causal pathway between NEFA and ISI. DAGitty recommended a minimal adjustment of waist, sex, and MET (physical activity).")
sfig_dag_issi2 <- sfig("dag_issi2", "Directed acyclic graph taken from DAGitty (DAGitty.net) for the potential causal pathway between NEFA and ISSI-2. DAGitty recommended a minimal adjustment of waist, ethnicity, and MET (physical activity).")
sfig_corr_conc <- sfig('corr_conc', 'Pearson correlation heatmap of non-esterified fatty acids (*nmol/mL*) and basic PROMISE participant characteristics for the baseline visit (2004-2006). Darkness of the colour indicates the magnitude of the correlation, with orange indicating positive and blue indicating negative correlations.')
sfig_corr_pct <- sfig('corr_pct', 'Pearson correlation heatmap of non-esterified fatty acids (*mol%*) and basic PROMISE participant characteristics for the baseline visit (2004-2006). Darkness of the colour indicates the magnitude of the correlation, with orange indicating positive and blue indicating negative correlations.')
```

@@

>>>>

sfig_consort <- sfig('consort', 'CONSORT diagram of sample size at each examination visit for the main PROMISE cohort in addition of the sample size determination for the present analysis.')
sfig_dag_isi <- sfig("dag_isi", "Directed acyclic graph taken from DAGitty (DAGitty.net) for the potential causal pathway between NEFA and ISI. DAGitty recommended a minimal adjustment of waist, sex, and MET (physical activity).")
sfig_dag_issi2 <- sfig("dag_issi2", "Directed acyclic graph taken from DAGitty (DAGitty.net) for the potential causal pathway between NEFA and ISSI-2. DAGitty recommended a minimal adjustment of waist, ethnicity, and MET (physical activity).")
sfig_corr_conc <- sfig('corr_conc', 'Pearson correlation heatmap of non-esterified fatty acids (*nmol/mL*) and basic PROMISE participant characteristics for the baseline visit (2004-2006). Darkness of the colour indicates the magnitude of the correlation, with orange indicating positive and blue indicating negative correlations.')
sfig_corr_pct <- sfig('corr_pct', 'Pearson correlation heatmap of non-esterified fatty acids (*mol%*) and basic PROMISE participant characteristics for the baseline visit (2004-2006). Darkness of the colour indicates the magnitude of the correlation, with orange indicating positive and blue indicating negative correlations.')
sfig_gee_adj_nowaist <- sfig('gee_adj_nowaist', 'Longitudinal associations of individual non-esterified fatty acids (mol% and nmol/mL) with insulin sensitivity and beta-cell function over the 6 years in the PROMISE cohort. Generalized estimating equation models were fully adjusted, *excluding waist circumference*. X-axis values represent a percent difference in the outcome per SD increase in the fatty acid. P-values were adjusted for the BH false discovery rate, with each increase in dot size and blackness representing a p-value significance of p>0.05, p<0.05, and p<0.01.')
```

@@

```{r ESM_Fig6, fig.cap=sfig_gee_adj_nowaist}
gee_results %>% 
    dplyr::filter(model == "Adjusted no waist") %>% 
    plot_gee_main()
```
